---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(rhxl)
```


# rhxl
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/dirkschumacher/rhxl?branch=master&svg=true)](https://ci.appveyor.com/project/dirkschumacher/rhxl)
[![Travis-CI Build Status](https://travis-ci.org/dirkschumacher/rhxl.svg?branch=master)](https://travis-ci.org/dirkschumacher/rhxl)
[![Coverage Status](https://img.shields.io/codecov/c/github/dirkschumacher/rhxl/master.svg)](https://codecov.io/github/dirkschumacher/rhxl?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rhxl)](https://cran.r-project.org/package=rhxl)

Humanitarian Exchange Language in R (work in progress)

The goal of this package is to provide functions to handle datasets with HXL tags. It is currently work in progress. If you have any ideas on how to further develop the package, please feel free to open an issue or a pull request.

> HXL is a different kind of data standard, designed to improve information sharing during a humanitarian crisis without adding extra reporting burdens.
[hxlstandard.org](http://hxlstandard.org/standard/1_0final/)

## Install

To install the current development version use devtools:

```R 
devtools::install_github("dirkschumacher/rhxl")
```

## Example

Read in the aiports of Viet Nam.

```{r}
data_url <- "http://ourairports.com/countries/VN/airports.hxl"
hxl_data <- as_hxl(read.csv(data_url))
head(hxl_data)
```


To get the schema:

```{r}
schema(hxl_data)
```

We can also test, if a dataset supports a schema. For example, we could test if the dataset has lat/lng coordinates.

```{r}
validate(hxl_data, c("#geo +lat", "#geo +lon"))
```

With this information you could for example write a function that can automatically display a dataset on a map - without much configuration by the user. It also makes sharing of information easier.

## API

Currently work in progress.

### as_hxl

`as_hxl`converts an object to an HXL data_frame. It returns a `tibble` with additional meta data for each column. 

```R
data <- as_hxl(readr::read_csv("treatment_centers.csv"))
```

### schema

Returns an informative object describing the schema of the dataset. Possibly a data_frame. 

Currently it returns a simple character vector.

```R
schema(hxl_table)
```


### validate

Validate an HXL data_frame against a schema. Returns a boolean if it matches the schema.

Returns TRUE/FALSE if a table matches a schema.

```R
validate(hxl_table, c("#adm1", "#adm2", "#affected"))
```

